---
title: "Example_scTypeGSEA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_scTypeGSEA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Description

This package is used to label cell type for each cell in single-cell data. The first step is to use “Seurat” package to do data pre-process and cells cluster. After clustering the cell, our purpose is to label cell type for each cluster. The second step is to do use differential gene expression analysis to find marker genes for each cluster, that is to find which genes have higher expression in given cluster and lower expression in others. Then for each cluster we get a list of marker genes. It is also known that each cell type has its own marker genes, thus the final step is to use the Gene set enrichment analysis (GSEA) method to compare two list of genes. For each cluster, we compare its marker genes to all cell type marker genes. GSEA will give us which cell type marker genes are the most relevant to such cluster marker genes. That is the cell type for such cluster.This package aims to assemble all these steps to give cell type for each cluster.

## Installation

You can use following codes to install the package.

```{r, eval=FALSE}
library(devtools)
install_github("cailab-tamu/scTypeGSEA")
library(scTypeGSEA)
```

## Example

One can do quality control, data pre-process, cluster, get full changes, do GSEA and label the cell in one step. Following code is for toy data set "pbmc_small".
```{r, warning=FALSE}
# pbmc_res <- assignCellType(obj = pbmc_small, min.cells = 1, min.features = 10, nfeatures = 100, npcs = 10, dims = 1:10, k.param = 5, resolution = 0.75, min.pct = 0.25, test.use = "MAST", minSize = 5)
```

It will return a list. The first entry is a Seurat object.
```{r}
pbmc_res$obj
```

The second entry is a matrix contains cluster-ID, cell type and p-value for each cell.
```{r}
head(pbmc_res$cell_mat)
```

Next is to illustrate step by step.

### Load data and quality control
Here we will use data set "pbmc" go through this pipline. First is to load data set and do quality control. The input can be a Seurat object or count gene expression matrix.
```{r, warning=FALSE}
library(scTypeGSEA)
pbmc <- scqc(pbmc_raw, min.cells = 3, min.features = 200, percent.mt = 5)
pbmc
```

### Data pre-process and do cluster
Next to do data pre-process and cluster. Here data pre-process includes normalization, feature selection and dimension reduction. Here we do “LogNormalize” nomalization, "vst" feature selections to select 2000 features and do PCA to get first 50 PCs. After data-process, we cluster the data.
```{r, warning=FALSE}
set.seed(47)
pbmc <- doClustering(pbmc, normalization.method = "LogNormalize", scale.factor = 10000, selection.method = "vst", nfeatures = 2000, npcs = 50, dims = 1:50, k.param = 20, resolution = 0.5)
head(pbmc@meta.data$seurat_clusters)
```

If your data set have already been clustered, you can set "doit = TRUE" to do cluster using new parameters.
```{r, eval=FALSE}
## don't run
pbmc <- check_cluster(pbmc, k.param = 30, resolution = 0.5, doit = TRUE)
```

### Get full changes
To get rank of gene list, we need to do gene differential expression analysis for each cluster with respect to all others. Here we use "wilcox" method.
```{r, warning=FALSE}
cluster_list <- getFC(pbmc, min.pct = 0.25, test.use = "wilcox")
head(cluster_list[[1]])
```

### Gene set enrichment analysis (GSEA)
After getting rank of gene list for each cluster, GSEA method will give us NSF. By NSF, we can decide the cell type for each cluster. Here we use "PanglaoDB" database.
```{r, warning=FALSE}
cluster_celltype <- doGSEA(cluster_list = cluster_list, db = "PanglaoDB_list",
                                  minSize = 15, maxSize = 500)
head(cluster_celltype)
```

You can also use your own cell type database. Here "otherdb" can be a path to the new data base that hope to be used, which is list of cell types with their marker genes. The file must be 'rds' format.
```{r, eval=FALSE}
## Don't run
cluster_celltype <- GSEA_analysis(cluster_list = cluster_list, otherdb = "path/to/your/rdsfile")
```

### Label the cell type
The last step is to add label cell type to our Seurat object.
```{r, warning=FALSE}
pbmc_res <- labelSeurat(pbmc, cluster_celltype)
pbmc_res$obj
head(pbmc_res$cell_mat)
```

### Plot the data with cell type

```{r, warning=FALSE}
pbmc <- pbmc_res$obj
pbmc <- Seurat::RunTSNE(pbmc, dims = 1:30)
Seurat::DimPlot(pbmc, reduction = "tsne", label = TRUE, pt.size = 0.5) + NoLegend()
```
